\documentclass{report}

\usepackage{dirtree}
\usepackage[dvipsnames]{xcolor}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algpseudocode}
\usepackage{todonotes}

\title{Optimization of Invariant Verification for Safe Replication of BFT-Logs}
\author{Pascal von Fellenberg}
\begin{document}
\maketitle
%\newcommand{\todo}[1]{\emph{\colorbox{Yellow}{TODO #1}}}
\newcommand{\balance}[0]{\textit{balance}}
\newcommand{\apply}[0]{\textit{apply}}
\newcommand{\acc}[0]{\textit{acc}}
\newcommand{\deltacc}[1]{#1^\delta}
\newcommand{\id}[1]{#1_\textit{id}}

\section*{Abstract}
TODO

\section{Introduction}
TODO

\section{Background / Terminology} \todo{add description of }
TODO

\section{Design of Test Repository}

\begin{tabular}{|l|l|}
\hline
\textbf{Git Commit Field} & \textbf{Purpose} \\
\hline
\emph{author} & $M_{author}$ as \emph{author.name}, the \emph{author.email} is not used; \\
\emph{commiter} & Not used; \\
\emph{parents} & First parent is always $M_{prev}$, while the next are $M_{deps}$; \\
\emph{message} & Contains both the $M_{payload}$ and $M_{signature}$; \\
\emph{tree-hash} & Hash referencing the tree-object that represents the delta state. \\
\hline
\end{tabular}

\subsection{State of the 2P-BFT-Log to Track}
idk what to write here yet.

\subsection{State of the $\delta$-GOC-Ledger to Track}
Each individual state represents a state of an account.
Each account stores 5 values: $A_{id}$, $A_{\uparrow}$, $A_{\downarrow}$, $A_{\rightarrow}$ and $A_{\leftarrow}$. $A_{id}$ is the account identifier and doesn't change.
$A_{\uparrow}$ and $A_{\downarrow}$ are numbers and represent the number of created and the number of destroyed tokens respectively.
$A_{\rightarrow}$ and $A_{\leftarrow}$ are mappings from account identifiers to numbers and represent the number of tokens given to and received (and acknowledged) from the author respectively.
All numbers in this account are grow-only counters, meaning their value cannot decrease over time.

To save space, only the attributes that changed since the last state are stored in the delta state, denoted as $A^\delta$.

A Ledger is a dictionary of accounts. It maps account identifiers to account states.

\subsection{Tree Objects as Delta-states}
\label{sec:tree-obj-desc}
Each delta state is encoded into a tree object.
This works as follows: The tree contains all the attributes that changed as entries.
The entries corresponding to $A_{\uparrow}$ and $A_{\downarrow}$ contain hashes to blob objects containing their respective number (number of created and destroyed tokens).
The entries corresponding to $A_{\leftarrow}$, $A_{\rightarrow}$ contain hashes to tree objects.
These tree objects contain entries corresponding to the authors which contain hashes to blob objects containing their respective number (number of given and acknowledged tokens).

\subsection{Git Repository as Ledger}
\begin{figure}[h!]
    \dirtree{%
    .1 refs/.
    .2 heads/.
    .3 frontier/.
    .4 <token type>/.
    .5 <author id>.
    .5 ....
    .4 ....
    }
    \caption{References directory in Git repository.}
\end{figure}
Since this is only a benchmark of the verification of commits, I don't include the references necessary for replication (remotes), the ones for checkpointing and the ones related to aliases.

\section{Invariants to Check}

\subsection{Invariants on delta account $\deltacc{A}$}
\label{sec:invariants-delta-a}
The following invariants depend on the state of the ledger before the update of $\deltacc A$ is performed ($L_\textit{before}$).
$L_\textit{after}$ represents the ledger after $\deltacc A$ has been applied to $L_\textit{before}$.
\begin{enumerate}
    \item [$\delta$1] (minimal) Any field in the delta account must increase the state of the Ledger (otherwise it would not have to be stored) so $L_\textit{before} \sqsubset L_\textit{after}$.
    \item [$\delta$2] (non-negative balance after giving) If $\deltacc A_\rightarrow$ is non-empty, the constraint $\balance(L_\textit{after}[\id{\deltacc{A}}]) \ge 0$ must hold.
    \item [$\delta$3] (non-negative balance after destroying) If $\deltacc A_\downarrow$ is present, the constraint $L_\textit{after}[\id{\deltacc A}] \ge 0$ must hold.
    \item [$\delta$4] (valid acknowledgements) If $\deltacc A_\leftarrow$ is non-empty, the constraint $\deltacc A_\leftarrow[x] \le (L_\textit{before}[x])_\rightarrow[\id{\deltacc A}]$ must hold for all $x \in \deltacc A_{\leftarrow*}$ ($\deltacc A_{\leftarrow*}$ is the set of keys in $\deltacc A_{\leftarrow}$, i.e. the author ids of authors from which the author with id $A_\textit{id}$ acked something).
    \item [$\delta$5] (valid creator) If $\deltacc A_\uparrow$ is present, $A$ must be permitted to create tokens. \todo{who can create tokens and when?}
\end{enumerate}

\subsection{Invariants on Message $M$ and corresponding delta account $\acc(M)$}
A message $M'$ is relevant to message $M$ if one of the following hold
\begin{itemize}
    \item $\acc(M)_{\rightarrow *}$ contains $M'_\textit{id}$ \todo{I am not sure if this dependency is necessary.. }
    \item $\acc(M)_{\leftarrow *}$ contains $M'_\textit{id}$
\end{itemize}

\textbf{Relevant Dependencies} \\
The dependencies of a message $M$ with corresponding delta account $\acc(M)$ are relevant if the following holds: For all $msgId \in M_\textit{deps}$, there exists $M'$ with $\texttt{MsgId}(M') = msgId$ and $M'$ is relevant to $M$.\\
In order for a message $M$ to be GOC-Ledger valid, the following must hold \footnote{in these invariants, we allow that a message has multiple dependencies from the same author. This is covered by the 2P-BFT-Log invariant M5.} \todo{Find a better name for this than ``GOC-Ledger valid''}
\begin{itemize}
    \item the invariants $\delta1$ to $\delta6$ (see Section \ref{sec:invariants-delta-a}) must be satisfied for $\acc(M)$
    \item (non-decreasing dates) $(M_\textit{prev})_\textit{date} \leq M_\textit{date}$ must hold
    \item the dependencies of $M$ are relevant
    \item for all $M' \in M_\textit{deps}\cup \{M_prev\}\}$, $M'$ must also be GOC-Ledger valid
\end{itemize}

\subsection{Invariants on Message $M$ inherited from 2P-BFT-Log}
\begin{enumerate}
    \item [M1] \emph{(single previous dependency)}: either $M_\textit{prev} = \bot$ or $M_\textit{prev} = \texttt{MsgId}(M'): M' \text{exists and is valid.}$
    \item [M2] \emph{(single author)}: if $M_\textit{prev} = \texttt{MsgId}(M')$ then $M'_\textit{author} = M_\textit{author}$.
    \item [M3] \emph{(valid external dependencies)}: for all $\textit{msgId} \in M_\textit{deps}$, there exists a valid $M'$ such that $\textit{MsgId}(M') = \textit{msgId}$ and $M'_\textit{author} \neq M_\textit{author}$.
    \item [M4] \emph{(single author dependencies)}: for any possible \textit{author} not equal to $M_\textit{author}$, there exists at most one $\textit{msgId} \in M_\textit{deps}$, such that $\texttt{MsgId}(M') = \textit{msgId}$ and $M'_\textit{author} = \textit{author}$.
    \item [M5] \emph{(self-certifying)}: $M_\textit{signature}$ is consistent with $M_\textit{author}$.
    \item [M6] \emph{(acyclic)}: $M_\textit{prev}$ does not reference a successor of $M$, i.e. ($M_\textit{prev} \neq M' : M \stackrel{\textit{log}}{\rightsquigarrow} M'$).
\end{enumerate}
Validity properties if log $L$ is in growing phase:
\begin{enumerate}
    \item [CL1] \emph{(no forks)}: $L_\textit{forks} = \emptyset$.
    \item [CL2] \emph{(valid last message)}: if $L_\textit{last} = M \neq \bot$ then $M$ is valid.
    \item [CL3] \emph{(consistent previous author)}: if $L_\textit{last} \neq \bot$ then $L_\textit{author} = (L_\textit{last})_\textit{author}$.
\end{enumerate}
Validity properties if log $L$ is in shrinking phase:
\begin{enumerate}
    \item [FL1] \emph{(non-empty forks)}: $L_\textit{forks} \neq \emptyset$.
    \item [FL2] \emph{(valid last message)}: same as CL2.
    \item [FL3] \emph{(consistent previouse author)}: same as CL3.
    \item [FL4] \emph{(valid forks)}: for all $M \in L_\textit{forks}$, $M$ is valid.
    \item [FL5] \emph{(consistent fork author)}: for all $M \in L_\textit{forks}$, $L_\textit{author} = M_\textit{author}$ and if $L_\textit{last} \neq \bot$ then $L_\textit{author} = (L_\textit{last})_\textit{author}$.
    \item [FL6] \emph{(valid proof)}: if $M \in L_\textit{forks}$ then $\exists M' \in L_\textit{forks}$ such that $M \neq M'$ and $M_\textit{prev} = M'_\textit{prev}$.
    \item [FL7] \emph{(consistent proof)}: for all $M \in L_\textit{forks}$, $L_\textit{last} = M_\textit{prev}$.
\end{enumerate}


\subsection{Validity Invariants Concerning Git Objects}
\begin{itemize}
    \item all the referenced git objects must exist
    \item the tree objects must be of the structure described in Section \ref{sec:tree-obj-desc}
    \item committer and author must be equal\footnote{therefore when a property must hold for the author, it must also hold for the committer}
    \item blob objects representing numbers must contain integers as bytes, with no leading zero bytes
    \item email of author must be of format \texttt{<id>@gitgen.com}
    \item id in the email must match the name of the author
    \item ...
\end{itemize}

\section{Verification Algorithm (GOC-Ledger Invariants)}
\begin{algorithmic}[1]
\Function{verify\_log}{set of Log Messages: $\mathcal M$}
\State $S \gets \emptyset$ \Comment{This is a dictionary with default value $\emptyset$ on access}
\State $R \gets \langle\rangle$ \Comment{This variable will contain the ids of all valid messages}
\ForAll{$M \in \mathcal{M}$ in any topological order}
    \State $L_\textit{before} \gets \bigsqcup_{L \in S[\id{M}]} L$
    \State $S[\id{M}] \gets \emptyset$
    \State $L_\textit{after} \gets \Call{apply}{M_\textit{acc}, L_\textit{before}}$
    \If{$\Call{account\_valid}{L_\textit{before}, L_\textit{after}, M}$}
        \State $R \gets R \| \langle M \rangle$
    \EndIf
    \State $L_M \gets L_\textit{after}$
    \ForAll{$\id{C} \in \Call{children\_ids}{M}$}
        \State $S[\id{C}] \gets S[\id{C}] \cup \{L_M\}$
    \EndFor
\EndFor
\Return R
\EndFunction
\end{algorithmic}
\todo{Implement APPLY function in pseudocode}
\todo{borrow notation from Heisch's Master Thesis}




\end{document}